---
title: "Statistics for Data Analytics"
author: "Anish Rao - 20066423"
output:
  word_document: default
  pdf_document: default
---
# Question 1 
## a) i) State the null and alternative hypothesis

Null Hypothesis (H0) - mean of petal length <= 3.5cm

Alternate Hypothesis (H1) - mean of petal length > 3.5cm

## a) ii) Compute t statistc value
```{r}
petal_length = c(4.5, 3.2, 5.1, 4.8, 6.3)
ttest = t.test(petal_length, mu = 3.5, alternative = 'greater')
cat('T-statistic value: \n')
ttest$statistic
```

## a) iii) Find the critical value
```{r}
a = 0.05
dof = 4
critical_val = qt(1- a, dof)
cat('Critical value:', critical_val)
```

## a) iv) Identify decision rule and express your decision
```{r}
if (ttest$statistic > critical_val){
  cat('Since T statistic (2.563078) > Critical value (2.131847), we reject H0')
  cat('\nThe Mean petal length is greater than 3.5 cm')
} else{
  cat("Since T statistic (2.563078) < Critical value (2.131847), we fail to reject H0")
}
```

## a) v) Compute p-value for above test and confirm your decision
```{r}
petal_length = c(4.5, 3.2, 5.1, 4.8, 6.3)
ttest = t.test(petal_length, mu = 3.5, alternative = 'greater')
cat('p value: \n')
ttest$p.value

if (ttest$p.value < 0.05){
  cat('Since p value (0.03121716) < significance level (0.05), we reject H0')
  cat('\nThe Mean petal length is greater than 3.5 cm')
} else{
  cat("Since p value (0.03121716) > significance level (0.05), we fail to reject H0")
}
```
--------------------------------------------------------------------

## b) i) State the null and alternative hypothesis

Null Hypothesis (H0) - variance of petal length = variance of petal width

Alternate Hypothesis (H1) - variance of petal length not equal to variance of petal width

## b) ii) Compute t statistc value
```{r}
petal_length = c(4.5, 3.2, 5.1, 4.8, 6.3)
petal_width = c(1.8, 1.3, 2.1, 1.9, 2.5)

cat('F statistic value: \n')
var.test(petal_length, petal_width)
```

## b) iii) Find the critical value
```{r}
a = 0.05
upper = qf(1 - a/2, df1 = 4, df2 = 4)
lower = qf(a/2, df1 = 4, df2 = 4)
cat('Critical value- \n')
cat("[", lower, ":", upper, "]")
```

## b) iv) Identify decision rule and express your decision
```{r}
Fstat = 6.4948
if (Fstat < lower || Fstat > upper){
  cat('Since F statistic is not in Critical value range, we reject H0')
} else{
  cat("F statistic is between critical value range, we fail to reject H0")
  cat("\nThere is not enough evidence to say variance are different for petal length and petal width")
}
```

## b) v) Can we conclude that different categories of "Species" have same probability to be observed at confidence level 0.90. Demonstrate your answer with an appropriate hypothesis test

```{r}
species = c("Setosa","Setosa", "Versicol", "Versicol", "Virginica")
observ_table = table(species)
probs = rep(1/3,3)
chi_test = chisq.test(observ_table, p = probs)

cat('Chi square test: \n')
chi_test
```
Null Hypothesis (H0) - The three species have same probability

Alternate Hypothesis (H1) - Atleast one species has different probability

Since the p-value (0.8187) > significance level (0.10), we fail to reject H0

Therefore, we can say all species may have same probability at confidence 90%.