---
title: "Statistics for Data Analytics"
author: "Anish Rao - 20066423"
date: "2025-05-08"
output: word_document
---
# Question 4
## i) Compute the median of variable length (feet) and interpret it.
```{r}
lengths = c(35, 22, 48, 10, 30, 28, 40, 32, 18, 45)
med_length = median(lengths)
cat("Median: \n")
med_length
```

## ii) Compute the IQR and standard deviation of variable length (feet).
```{r}
iqr_length = IQR(lengths)
cat("IQR: \n")
iqr_length

sd_length = sd(lengths)
cat("Standard Deviation: \n")
sd_length
```

## iii) Using chebyshevs rule find the three-sigma interval for length (feet).
```{r}
mean_length = mean(lengths)

lower = mean_length - 3*sd_length
upper = mean_length + 3*sd_length
cat("Chebyshev Interval: \n")
cat("[", lower, ":", upper, "]")
```

## iv) Using boxplot technique, specify outliers for length (feet), if any.
```{r}
boxplot_length = boxplot.stats(lengths)
cat("Outliers by Boxplot technique: \n")
print(boxplot_length$out)
cat("No outliers are detected by the boxplot rule")
```

## v) Use (c) and (d) to propose a minimal interval estimation for outlier detection for length (feet).

There are no outliers according to boxplot technique while the chebyshev interval includes all values so proposed range would be the dataset range [10, 48]. 
